\texttt{thornado}-transport: IMEX schemes for two-moment neutrino transport respecting Fermi-Dirac statistics

### 2019-01-25 Documents/M1/Astronum_2018/manuscript.tex
Work on Astronum paper.

\author{Ran Chu}
\address{Department of Physics and Astronomy, University of Tennessee Knoxville, TN 37996-1200}
\ead{rchu@vols.utk.edu}

\author{Eirik Endeve}
\address{Computational and Applied Mathematics Group, Oak Ridge National Laboratory, Oak Ridge, TN 37831 USA}
\address{Department of Physics and Astronomy, University of Tennessee Knoxville, TN 37996-1200}
\ead{endevee@ornl.gov}

\author{Cory D. Hauck}
\address{Computational and Applied Mathematics Group, Oak Ridge National Laboratory, Oak Ridge, TN 37831 USA}
\address{Department of Mathematics, University of Tennessee Knoxville, TN 37996-1320}
\ead{hauckc@ornl.gov}

\author{Anthony Mezzacappa}
\address{Department of Physics and Astronomy, University of Tennessee Knoxville, TN 37996-1200}
\ead{mezz@utk.edu}

\author{Bronson Messer}
\address{Scientific Computing and Theoretical Physics Groups, Oak Ridge National Laboratory, Oak Ridge, TN 37831 USA}
\address{Department of Physics and Astronomy, University of Tennessee Knoxville, TN 37996-1200}
\ead{bronson@ornl.gov}

\begin{abstract}
We develop implicit-explicit (IMEX) schemes for neutrino transport through a background material in the \texttt{thornado} code in the context of a two-moment model that evolves the angular moments of a neutrino phase-space distribution function.
Considering the upper and lower bounds that are introduced by Pauli's exclusion principle on the moments, an algebraic moment closure based on Fermi-Dirac statistics and a convex-invariant time integrator both are demanded.
A finite-volume/first-order discontinuous Galerkin(DG) method is used to interpret how an algebraic moment closure based on Fermi-Dirac statistics satisfies the bounds.
Several algebraic closures are compared with these bounds in mind and the Cernohorsky and Bludman closure\cite{cernohorskyBludman_1994}, which satisfies the bounds, is chosen for our IMEX schemes.
For the convex-invariant time integrator, two IMEX schemes named PD-ARS have been proposed following the work in \cite{chu_etal_2018}.
PD-ARS denotes a convex-invariant IMEX Runge-Kutta scheme that has the first-order accuracy in general, high-order accuracy in the streaming limit, and works well in the diffusion limit.
Our two PD-ARS schemes are second- and third-order, explicit, strong-stability-preserving Runge-Kutta methods\cite{gottlieb_etal_2001} as their explicit part, respectively, and therefore are second- and third-order accurate in the streaming limit, respectively.
The accuracy and convex-invariance of our PD-ARS schemes are demonstrated in the numerical tests with a third-order DG method for spatial discretization and a simple Lax-Friedrichs flux.
We show preliminary result employing tabulated neutrino opacities from Bruenn(1985)\citep{Bruenn_1985}

>>>>

\author{Ran Chu}
\address{Department of Physics and Astronomy, University of Tennessee, Knoxville, TN 37996-1200}
\ead{rchu@vols.utk.edu}

\author{Eirik Endeve}
\address{Computational and Applied Mathematics Group, Oak Ridge National Laboratory, Oak Ridge, TN 37831 USA}
\address{Department of Physics and Astronomy, University of Tennessee, Knoxville, TN 37996-1200}
\address{Joint Institute for Computational Sciences, Oak Ridge National Laboratory, Oak Ridge, TN 37831 USA}
\ead{endevee@ornl.gov}

\author{Cory D. Hauck}
\address{Computational and Applied Mathematics Group, Oak Ridge National Laboratory, Oak Ridge, TN 37831 USA}
\address{Department of Mathematics, University of Tennessee, Knoxville, TN 37996-1320}
\ead{hauckc@ornl.gov}

\author{Anthony Mezzacappa}
\address{Department of Physics and Astronomy, University of Tennessee, Knoxville, TN 37996-1200}
\address{Joint Institute for Computational Sciences, Oak Ridge National Laboratory, Oak Ridge, TN 37831 USA}
\ead{mezz@utk.edu}

\author{Bronson Messer}
\address{Scientific Computing and Theoretical Physics Groups, Oak Ridge National Laboratory, Oak Ridge, TN 37831 USA}
\address{Department of Physics and Astronomy, University of Tennessee, Knoxville, TN 37996-1200}
\ead{bronson@ornl.gov}

\begin{abstract}
We develop implicit-explicit (IMEX) schemes for neutrino transport in a background material using the \texttt{thornado} code in the context of a two-moment model that evolves the angular moments of a neutrino phase-space distribution function.
Considering the upper and lower bounds that are introduced by Pauli's exclusion principle on the moments, an algebraic moment closure based on Fermi-Dirac statistics and a convex-invariant time integrator both are demanded.
A finite-volume/first-order discontinuous Galerkin(DG) method is used to interpret how an algebraic moment closure based on Fermi-Dirac statistics satisfies the bounds.
Several algebraic closures are compared with these bounds in mind and the Cernohorsky and Bludman closure\cite{cernohorskyBludman_1994}, which satisfies the bounds, is chosen for our IMEX schemes.
For the convex-invariant time integrator, two IMEX schemes named PD-ARS have been proposed following the work in \cite{chu_etal_2018}.
PD-ARS denotes a convex-invariant IMEX Runge-Kutta scheme that has first-order accuracy in general, high-order accuracy in the streaming limit, and works well in the diffusion limit.
Our two PD-ARS schemes are second- and third-order, explicit, strong-stability-preserving Runge-Kutta methods\cite{gottlieb_etal_2001} in their explicit part, respectively, and therefore are second- and third-order accurate in the streaming limit, respectively.
The accuracy and convex-invariance of our PD-ARS schemes are demonstrated in the numerical tests with a third-order DG method for spatial discretization and a simple Lax-Friedrichs flux.
We show preliminary result employing tabulated neutrino opacities. 
%from Bruenn(1985)\citep{Bruenn_1985}

### 2019-01-25 Documents/M1/Astronum_2018/sections/Acknowledge.tex
Work on astronum paper
Work on Astronum paper.

This research is sponsored, in part, by the Laboratory Directed Research and Development Program of Oak Ridge National Laboratory (ORNL), managed by UT-Battelle, LLC for the U. S. Department of Energy under Contract No. De-AC05-00OR22725.  
This research was supported by the Exascale Computing Project (17-SC-20-SC), a collaborative effort of the U.S. Department of Energy Office of Science and the National Nuclear Security Administration.  
This material is based, in part, upon work supported by the U.S. Department of Energy, Office of Science, Office of Advanced Scientific Computing Research.
Eirik Endeve and Anthony Mezzacappa acknowledge support from the National Science Foundation under group NSF-GP-... and from the UT-ORNL Joint Institute for Computational Sciences.

>>>>

This research is sponsored, in part, by the Laboratory Directed Research and Development Program of Oak Ridge National Laboratory (ORNL), managed by UT-Battelle, LLC for the U. S. Department of Energy under Contract No. De-AC05-00OR22725. This material is based, in part, upon work supported by the U.S. Department of Energy, Office of Science, Office of Advanced Scientific Computing Research. This research was also supported by the Exascale Computing Project (17-SC-20-SC), a collaborative effort of the U.S. Department of Energy Office of Science and the National Nuclear Security Administration, and by the National Science Foundation Gravitational Physics Program (NSF-GP 1505933 and 1806692).

### 2019-01-25 Documents/M1/Astronum_2018/sections/Numerical.tex
Work on Astronum paper.

In this section we present numerical results obtained with the PD-ARS schemes given in Section~\ref{se:TimeIntegration}.
The tests in Section \ref{se: Accuracy Tests} are designed to compare the accuracy of the schemes in streaming, absorption, and scattering-dominated regimes in one spatial dimension.
The test in Section \ref{se: Neutrino Stationary State Test} demonstrates the convex-invariance of PD-ARS schemes schemes.
All the tests in this section were computed with third-order accurate spatial discretization (polynomials of degree $k=2$) and time step $\dt = 0.1 \times \dx $, using the DG scheme from \cite{chu_etal_2018}.

@@

To compare the accuracy of the IMEX schemes, we applied our PD-ARS schemes and the SSP2332 scheme from Pareschi \& Russo \cite{pareschiRusso_2005} to problems with known smooth solutions in streaming, absorption(damping), and scattering-dominated(diffusing) regimes in one spatial dimension.  
All the tests in this subsection were computed with the maximum entropy closure in the low-occupancy limit (i.e., the Minerbo closure).  
In the streaming test, the second- and third-order accurate explicit strong-stability-preserving Runge-Kutta methods from \cite{gottlieb_etal_2001} (SSPRK2 and SSPRK3, respectively) are also included.  
To compare the numerical results to analytic solutions, the averaged absolute error or the averaged relative error are computed in the $L^{1}$-error norm: the absolute error for the streaming test and the diffusion test; the relative error for the damping test.
They are averaged over the cell with a equal weight quadrature for the cell integrals.
And to examine the convergence, we let the number of elements ($N$) vary from 8 to 128.

@@

The sine wave streaming test is designed to test accuracy in the free-steaming regime, i.e. $\sigma_{\Ab} = \sigma_{\Scatt} = 0$.
A periodic domain of unit length is used and the initial condition is $\cJ_{0} = \bcH_{0} = 0.5+0.49\times\sin\big(2\pi\,x\big)$.
We evolve the test until the sine wave has completed 10 crossings of the domain.
Figure~\ref{fig: SineWaveStreaming} plots the absolute error for the number density versus the number of elements $N$.
We see the errors obtained with SSPRK3 and PD-ARS with SSPRK3 decrease as $N^{-3}$, as expected.
All the other schemes have errors decreasing as $N^{-2}$.
Among the second-order schemes, SSP2332 has the smallest error.
In the streaming limit, the PD-ARS schemes limit to SSPRK2 and SSPRK3, respectively.
Therefore, the absolute errors of PD-ARS and SSPRK are identical.

\begin{figure}[h]
  \centering
    \includegraphics[width=0.8\textwidth]{figures/SineWaveStreaming}
   \caption{Absolute error versus number of elements $N$ for the streaming sine wave test.  Results employing various time stepping schemes are compared: SSPRK2 (cyan triangles pointing up), SSPRK3 (cyan triangles pointing down), SSP2332 (green crosses), PD-ARS with SSPRK2 (light red circles) and PD-ARS with SSPRK2 (light red hexagrams). Black dashed reference lines are proportional to $N^{-2}$ (top), and $N^{-3}$ (bottom), respectively.}
   \label{fig: SineWaveStreaming}
\end{figure}

@@

This test, adapted from \cite{skinnerOstriker_2013}, is designed for absorption-dominated regimes, with $\sigma_{\Scatt} = 0$ and $f_0 = 0$, which results in exponential damping of the wave amplitude.
A periodic domain of unit length and initial condition $\cJ_{0} = \bcH_{0} = 0.5+0.49\times\sin\big(2\pi\,x\big)$ is used.
The amplitude of the analytical solution decreases as $e^{-\sigma_{\Ab} t}$.
For $\sigma_{\Ab}$ = 0.1, 1 and 10 we evolve the test until the initial condition has been damped by a factor $e^{-10}$. 
Figure~\ref{fig:SineWaveDamping} shows convergence results of the sine wave damping test in the relative error.
SSP2332, a second-order scheme, displays second-order convergence rate.  
The PD-ARS schemes are only first-order accurate.  
\begin{figure}[h]
  \centering
    \includegraphics[width=0.9\textwidth]{figures/SineWaveDamping}
   \caption{Relative error versus number of elements, $N$, for the damping sine wave test. Results for different values of the absorption opacity $\sigma_{\Ab}$, employing various IMEX time stepping schemes, are compared.  Errors for $\sigma_{\Ab}=0.1$, $1$, and $10$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are SSP2332 ($+$), PD-ARS with SSPRK2 (circles) and PD-ARS with SSPRK3 (hexagram).  Black dashed reference lines are proportional to $N^{-1}$ (top) and $N^{-2}$ (bottom), respectively.}
  \label{fig:SineWaveDamping}
\end{figure}

@@

The last test with known smooth solution, adopted from \cite{radice_etal_2013}, is the sine wave diffusion test, i.e. $\sigma_{\Ab} = 0$ and $f_0 = 0$.
A periodic domain $D=\{x:x\in[-3,3]\}$ with initial conditions $\cJ_{0} = 0.5+0.49\times\sin\big(\f{\pi\,x}{3}\big)$ and $\cH_{0} =-\f{1}{3\sigma_{\Scatt}}\pderiv{\cJ_{0}}{x}$ are used.  
The reference diffusion solution is given by $\cJ = \cJ_{0} \times \exp\big(-\f{\pi^2 t}{27\sigma_{\Scatt}}\big)$ and $\cH = (3\,\sigma_{\Scatt})^{-1}\pd{\cJ}{x}$.  
We evolve with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$, and adjust the end time so that $t_{\mbox{\tiny end}}/\sigma_{\Scatt}=1$. 
Then the amplitude of the sine wave has been reduced by a factor $e^{-\pi^{2}/27}\approx0.694$ for all values of $\sigma_{\Scatt}$. 
Figure~\ref{fig:SineWaveDiffusionJ} shows the error versus $N$.  
SSP2332 and PD-ARS schemes display third-order accuracy for the number density, $\cJ$, and second-order accuracy for $\cH_{x}$, and their errors are difficult to distinguish.
PD-ARS with SSPRK2 behaves as good as SSP2332 in the diffusion region but requires one less implicit solves per time step.

\begin{figure}[h]
  \centering
  \centerline{\includegraphics[width=1.2\textwidth]{figures/SineWaveDiffusion}}
   \caption{Absolute error for the number density $\cJ$ (left) and the number flux $\cH_{x}$ (right) versus number of elements for the sine wave diffusion test.  Results with different values of the scattering opacity $\sigma_{\Scatt}$, employing different IMEX schemes, are compared.  Errors with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are:  SSP2332 ($+$), PD-ARS with SSPRK2 (circles), and PD-ARS with SSPRK3 (hexagram). Black dashed lines in the left plot are reference lines proportional to $N^{-2}$ (top) and $N^{-3}$ (bottom), respectively. Black dashed line in the right plot is a reference line proportional to $N^{-2}$ .}
   \label{fig:SineWaveDiffusionJ}
\end{figure}

@@

In this section we consider a more ``realistic'' test: two-dimensional multigroup neutrino transport with emission, absorption, and elastic scattering through a stationary background.  
This test is designed to test the realizability-preserving properties of the PD-ARS schemes.  
Figure~\ref{fig:NeutrinoStationaryTestEOS} plots the thermal state and the opacities used (adopted from \cite{Bruenn_1985}).  
This test is computed on a two-dimensional domain $D=\{\vect{x}\in\bbR^{2}:x^{1}\in[0,200], x^{2}\in[0,200]\}$ with a grid of 128 elements in each direction, a reflecting inner boundary and a homogeneous out-flowing outer boundary.
We use 10 energy groups to cover the range from 0~MeV to 300~MeV and initialize the neutrino number density to $\cJ = 10^{-99}$, flux density $\bcH=0$, and evolve until an approximate steady state is reached ($t=5$~ms).  
For this test, we employed the CB closure.  
We attempted to run this test with our PD-ARS schemes, SSP2332 from \cite{pareschiRusso_2005}, IMEXRKCB2 proposed by Cavaglieri \& Bewley \cite{cavaglieriBewley2015}, and the scheme proposed by McClarren et al. \cite{mcclarren_etal_2008}.
Only the PD-ARS schemes produce realizable moments and are able to evolve to a steady state.
The other schemes crashed after a few steps, when moments became non-realizable and the closure procedure failed.

Results obtained with the PD-ARS IMEX schemes are plotted in Figure~\ref{fig:NeutrinoStationaryTestEvolve} for various times: $t=0.01$~ms (top panels), 0.35~ms (middle panels), and 5.0~ms (bottom panels).  
In the left column we plot the solution in the $|\vect{\cH}|\cJ$-plane.  
In the middle column we show scatter plots of the number density $\cJ$ versus radius for select neutrino energies: 5~MeV (red lines), 16~MeV (magenta lines), and $93$~MeV (blue lines).  
In the right column we plot the flux factor $|\vect{\cH}|/\cJ$ versus radius for the same neutrino energies as in the middle column.  
In the left panels, each solution point in the domain is marked as a red dot in the $|\vect{\cH}|\cJ$-plane, and the realizable domain is shown as the light blue region.  
The figures show that all the states in the simulation with the PD-ARS schemes are realizable.

\begin{figure}[h]
  \centering
  \begin{tabular}{cc}
    \includegraphics[width=0.45\textwidth]{figures/NStatinaryS_EOS}
    \includegraphics[width=0.45\textwidth]{figures/NSS_Opacities}
  \end{tabular}
   \caption{Left panel: thermal state of the background versus radius in the neutrino stationary state test; mass density (solid line), temperature (dashed line), and electron fraction (dotted line).  Right panel: corresponding opacities for select neutrino energies; absorptivity ($\sigma_{\Ab}$) and scattering opacity ($ \sigma_{\Scatt}$).}
   \label{fig:NeutrinoStationaryTestEOS}
\end{figure}

\begin{figure}[h]
  \centering
    \includegraphics[width=\textwidth]{figures/NSS_1_1}\\
    \includegraphics[width=\textwidth]{figures/NSS_3_1} \\
    \includegraphics[width=\textwidth]{figures/NSS_5_1} \\
    \caption{Results from the neutrino stationary state test: moments relative to the realizable domain (left column, light blue domain is for $f \in [0,1]$ with black-solid-line as boundary, while light red is for $f\geq 0$ with thin-red-solid-line as boundary), the number density $\cJ$ versus radius (middle column), and the flux factor $|\bcH|/\cJ$ versus radius (right column), at $t = 0.01$~ms, 0.35~ms, and 5.0~ms.  For the plots in the left column, each $\bcM=(\cJ,\bcH)^{T}$ state is marked by a red dot, which are all inside the light blue region (the realizable domain for fermions).  The results of PD-ARS with SSPRK2 and PD-ARS with SSPRK3 are indistinguishable in these plots.}
    \label{fig:NeutrinoStationaryTestEvolve}
\end{figure}

>>>>

In this section we present numerical results obtained with the PD-ARS schemes given in Section~\ref{se:TimeIntegration}.
The tests in Section \ref{se: Accuracy Tests} are designed to compare the accuracy of the schemes in streaming, absorption, and scattering-dominated regimes in one spatial dimension.
The test in Section \ref{se: Neutrino Stationary State Test} demonstrates the convex-invariance of PD-ARS schemes.
All the tests in this section were computed with third-order accurate spatial discretization (polynomials of degree $k=2$) and time step $\dt = 0.1 \times \dx $, using the DG scheme from~\cite{chu_etal_2018}.

@@

To compare the accuracy of the IMEX schemes, we applied our PD-ARS schemes and the SSP2332 scheme from Pareschi \& Russo~\cite{pareschiRusso_2005} to problems with known smooth solutions in streaming, absorption(damping), and scattering-dominated(diffusing) regimes in one spatial dimension.  
All the tests in this subsection were computed with the maximum entropy closure in the low-occupancy limit (i.e., the Minerbo closure).  
In the streaming test, the second- and third-order accurate explicit strong-stability-preserving Runge-Kutta methods from~\cite{gottlieb_etal_2001} (SSPRK2 and SSPRK3, respectively) are also included.  
To compare the numerical results to analytic solutions, the averaged absolute error or the averaged relative error are computed in the $L^{1}$-error norm. We compute the absolute error for the streaming test and the diffusion test and the relative error for the damping test.
They are averaged over the cell with an equal-weight quadrature for the cell integrals.
And to examine the convergence, we let the number of elements ($N$) vary from 8 to 128.

@@

The sine wave streaming test is designed to test accuracy in the free-steaming regime; i.e. $\sigma_{\Ab} = \sigma_{\Scatt} = 0$.
A periodic domain of unit length is used and the initial condition is $\cJ_{0} = \bcH_{0} = 0.5+0.49\times\sin\big(2\pi\,x\big)$.
We evolve the test until the sine wave has completed 10 crossings of the domain.
Figure~\ref{fig: SineWaveStreaming} plots the absolute error for the number density versus the number of elements $N$.
We see the errors obtained with (SSPRK3) and (PD-ARS with SSPRK3) are smallest and decrease as $N^{-3}$, as expected for a scheme combining third-order accurate time stepping with third-order accurate spatial discretization.
For all the other schemes, using second-order accurate explicit time stepping, the error decreases as $N^{-2}$.
Among the second-order schemes, SSP2332 has the smallest error.
In the streaming limit, the PD-ARS schemes reduce to SSPRK schemes, such that (PD-ARS with SSPRK2) to SSPRK2 and (PD-ARS with SSPRK3) to SSPRK3, respectively.
Therefore, the absolute errors of PD-ARS schemes and SSPRK schemes are indistinguishable.

\begin{figure}[h]
  \centering
    \includegraphics[width=0.8\textwidth]{figures/SineWaveStreaming}
   \caption{Absolute error versus number of elements, $N$, for the streaming sine wave test.  Results employing various time stepping schemes are compared: SSPRK2 (cyan triangles pointing up), SSPRK3 (cyan triangles pointing down), SSP2332 (green crosses), PD-ARS with SSPRK2 (light red circles) and PD-ARS with SSPRK3 (light red asterisks). Black dashed reference lines are proportional to $N^{-2}$ (top), and $N^{-3}$ (bottom), respectively.}
   \label{fig: SineWaveStreaming}
\end{figure}

@@

The next test, adapted from~\cite{skinnerOstriker_2013}, is designed for absorption-dominated regimes, with $\sigma_{\Scatt} = 0$ and $f_0 = 0$, which results in exponential damping of the wave amplitude.
A periodic domain $D=\{x:x\in[0,1]\}$ and initial condition $\cJ_{0} = \bcH_{0} = 0.5+0.49\times\sin\big(2\pi\,x\big)$ is used.
The amplitude of the analytical solution decreases as $e^{-\sigma_{\Ab} t}$.
For $\sigma_{\Ab}$ = 0.1, 1 and 10 we evolve the test until the initial condition has been damped by a factor $e^{-10}$. 
Figure~\ref{fig:SineWaveDamping} shows convergence results of the test in the relative error.
Results for $\sigma_{\Ab}=0.1$, $1$, and $10$ are plotted with red, green, and blue lines, respectively.
SSP2332 is the most accurate among these schemes for $\sigma_{\Ab}= 1$ and 10. 
For $\sigma_{\Ab}=0.1$, PD-ARS with SSPRK2 is the most accurate schemes for $N=8$ and $N=16$.
Only SSP2332, a second-order accurate scheme, displays a second-order convergence rate.  
The PD-ARS schemes are first-order accurate in the damping test as expected.

\begin{figure}[h]
  \centering
    \includegraphics[width=0.9\textwidth]{figures/SineWaveDamping}
   \caption{Relative error versus number of elements, $N$, for the damping sine wave test. Results for different values of the absorption opacity $\sigma_{\Ab}$, employing various IMEX time stepping schemes, are compared.  Errors for $\sigma_{\Ab}=0.1$, $1$, and $10$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are SSP2332 ($+$), PD-ARS with SSPRK2 (triangles pointing up) and PD-ARS with SSPRK3 (triangles pointing down).  Black dashed reference lines are proportional to $N^{-1}$ (top) and $N^{-2}$ (bottom), respectively.}
  \label{fig:SineWaveDamping}
\end{figure}

@@

The final test with known smooth solution, adopted from~\cite{radice_etal_2013}, is the sine wave diffusion test, i.e. $\sigma_{\Ab} = 0$ and $f_0 = 0$.
A periodic domain $D=\{x:x\in[-3,3]\}$ with initial conditions $\cJ_{0} = 0.5+0.49\times\sin\big(\f{\pi\,x}{3}\big)$ and $\cH_{0} =-\f{1}{3\sigma_{\Scatt}}\pderiv{\cJ_{0}}{x}$ are used.  
The reference diffusion solution is given by $\cJ = \cJ_{0} \times \exp\big(-\f{\pi^2 t}{27\sigma_{\Scatt}}\big)$ and $\cH = (3\,\sigma_{\Scatt})^{-1}\pd{\cJ}{x}$.  
We evolve with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$, and adjust the end time so that $t_{\mbox{\tiny end}}/\sigma_{\Scatt}=1$, 
at which time the amplitude of the sine wave has been reduced by a factor $e^{-\pi^{2}/27}\approx0.694$ for all values of $\sigma_{\Scatt}$. 
Figure~\ref{fig:SineWaveDiffusionJ} shows the absolute error, obtained using different values of $\sigma_{\Scatt}$, for various IMEX schemes at $t=t_{\mbox{\tiny end}}$, versus $N$. 
Results for $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$ are plotted with red, green, and blue lines, respectively.
SSP2332 and PD-ARS schemes display third-order accuracy for the number density, $\cJ$, and second-order accuracy for $\cH_{x}$, and their errors are difficult to distinguish.
For $\sigma=10^{2}$, the errors on the number density $\cJ$ do not drop below $10^{-6}$ because of differences between the two-moment model and the diffusion equation used to obtain the analytic solution.
For larger values of the scattering opacity, $\sigma=10^{3}$ or $10^{4}$, the two-moment model agrees better with the diffusion model, and we observe convergence over the entire range of $N$.
PD-ARS with SSPRK2 behaves as well as SSP2332 in the diffusion region but requires 33\% less implicit solves per time step.

\begin{figure}[h]
  \centering
  \centerline{\includegraphics[width=1.2\textwidth]{figures/SineWaveDiffusion}}
   \caption{Absolute error for the number density $\cJ$ (left) and the number flux $\cH_{x}$ (right) versus number of elements for the sine wave diffusion test.  Results with different values of the scattering opacity, $\sigma_{\Scatt}$, employing different IMEX schemes, are compared.  Errors with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are:  SSP2332 ($+$), PD-ARS with SSPRK2 (circles), and PD-ARS with SSPRK3 (hexagram). Black dashed lines in the left plot are reference lines proportional to $N^{-2}$ (top) and $N^{-3}$ (bottom), respectively. Black dashed line in the right plot is a reference line proportional to $N^{-2}$ .}
   \label{fig:SineWaveDiffusionJ}
\end{figure}

@@

In this section we consider a more ``realistic'' test: two-dimensional Newtonian multigroup neutrino transport with emission, absorption, and isoenergetic scattering through a stationary background.  
This test is designed to test the realizability-preserving properties of the PD-ARS schemes.  
Figure~\ref{fig:NeutrinoStationaryTestEOS} plots the thermal state and the opacities used (adopted from~\cite{Bruenn_1985}).  
This test is computed on a two-dimensional domain $D=\{\vect{x}\in\bbR^{2}:x^{1}\in[0,200], x^{2}\in[0,200]\}$ with a grid of 128 elements in each direction, a reflecting inner boundary and a homogeneous out-flowing outer boundary.
We use 10 energy groups to cover the range from 0~MeV to 300~MeV and initialize the neutrino number density to $\cJ = 10^{-99}$, flux density $\bcH=0$, and evolve until an approximate steady state is reached ($t=5$~ms).  
For this test, we employed the CB closure.  
We attempted to run this test with our PD-ARS schemes, SSP2332 from~\cite{pareschiRusso_2005}, IMEXRKCB2 proposed by Cavaglieri \& Bewley~\cite{cavaglieriBewley2015}, and the scheme proposed by McClarren et al.~\cite{mcclarren_etal_2008}.
Only the PD-ARS schemes produce realizable moments and are able to evolve to a steady state.
The other schemes crashed after a few time steps, when moments became non-realizable and the closure procedure failed.

Results obtained with the PD-ARS IMEX schemes are plotted in Figure~\ref{fig:NeutrinoStationaryTestEvolve} for various times: $t=0.01$~ms (top panels), 0.35~ms (middle panels), and 5.0~ms (bottom panels).  
In the left column we plot the solution in the $|\vect{\cH}|\cJ$-plane.  
In the middle column we show scatter plots of the number density $\cJ$ versus radius for select neutrino energies: 5~MeV (red lines), 16~MeV (magenta lines), and $93$~MeV (blue lines).  
In the right column we plot the flux factor $|\vect{\cH}|/\cJ$ versus radius for the same neutrino energies as in the middle column. 
In the left panels, each solution point in the domain is marked as a red dot in the $|\vect{\cH}|\cJ$-plane, and the realizable domain is shown as the light blue region.  
The figures show that all the states in the simulation with the PD-ARS schemes are realizable.
In the middle column, we can see how the neutrinos are generated near the core, stream out, and reach a equilibrium distribution over the phase space eventually.
In the right column, we can see how the scattering reshapes the flux direction. 
The oscillatory behavior which is more obvious on high energy group is caused by the machine accuracy error.

\begin{figure}[h]
  \centering
  \begin{tabular}{cc}
    \includegraphics[width=0.45\textwidth]{figures/NStatinaryS_EOS}
    \hspace{30pt}
    \includegraphics[width=0.45\textwidth]{figures/NSS_Opacities}
  \end{tabular}
   \caption{Left panel: thermal state of the background versus radius in the neutrino stationary state test; mass density (solid line), temperature (dashed line), and electron fraction (dotted line).  Right panel: corresponding opacities for select neutrino energies; absorptivity ($\sigma_{\Ab}$) and scattering opacity ($ \sigma_{\Scatt}$).}
   \label{fig:NeutrinoStationaryTestEOS}
\end{figure}

\begin{figure}[h]
  \centering
    \includegraphics[width=\textwidth]{figures/NSS_1_1}\\
    \includegraphics[width=\textwidth]{figures/NSS_3_1} \\
    \includegraphics[width=\textwidth]{figures/NSS_5_1} \\
    \caption{Results from the neutrino stationary state test: moments relative to the realizable domain (left column, the light blue domain is for $f \in [0,1]$ with the black-solid-line as its boundary, while the light red domain is for $f\geq 0$ with the thin-red-solid-line as its boundary), the number density $\cJ$ versus radius (middle column), and the flux factor $|\bcH|/\cJ$ versus radius (right column), at $t = 0.01$~ms, 0.35~ms, and 5.0~ms.  For the plots in the left column, each $\bcM=(\cJ,\bcH)^{T}$ state is marked by a red dot, which are all inside the light blue region (the realizable domain for fermions).  The results of (PD-ARS with SSPRK2) and (PD-ARS with SSPRK3) are indistinguishable in these plots.}
    \label{fig:NeutrinoStationaryTestEvolve}
\end{figure}
